---
title: "svar oil"
output:
  html_document: default
  pdf_document: default
date: "2023-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SVAR oil to cpi

With the purpose to verify or refute the proposed hypothesis, will be incorporated methods statistics to analysis of the economic series to complement with technical content to the empirical content of the economic theories that explain the phenomenon raised.
For this, the use of a structural model of autoregressive vectors (sVAR) is proposed to analyze the behavior of inflation in Guatemala in the face of variations in the international prices of oil and its derivatives, given the importance that these have over all agents. from the final producers and consumers, to the way in which the state authorities formulate their economic policy.

```{r library, include=FALSE}
library(MASS)
library(strucchange)
library(zoo)
library(sandwich)
library(lmtest)
library(urca)
library(vars)
library(mFilter)
library(tseries)
library(forecast)
library(tidyverse)
library(ggplot2)
library(TSstudio)

library(readxl)
VAR_geometric_data <- read_excel("~/R/Oil to IPC/VAR_geometric_data.xlsx")
View(VAR_geometric_data)

loc_geo_y <- ts(VAR_geometric_data$local_price_yoy, start = c(2001,2), frequency = 12)
gas_geo_y <- ts(VAR_geometric_data$gas_yoy, start = c(2001,2), frequency = 12)
oil_geo_y <- ts(VAR_geometric_data$oil_yoy, start = c(2001,2), frequency = 12)
cpi_geo_y <- ts(VAR_geometric_data$inflation_yoy, start = c(2001,2), frequency = 12)
```
##Visualize data 
X axis: Time from 1M2001 to 12M2022
Y axis: Data Percentage %
```{r, echo = FALSE}
#Visualize data 

#ts_plot(cpi_geo_y)
#ts_plot(loc_geo_y)
#ts_plot(gas_geo_y)
#ts_plot(oil_geo_y)

autoplot(cbind(loc_geo_y, gas_geo_y, oil_geo_y, cpi_geo_y))

d <- ggplot(data = VAR_geometric_data)
d + geom_smooth(method = lm, mapping = aes(x = oil_yoy, y = inflation_yoy)) + geom_point(mapping = aes(x = oil_yoy, y = inflation_yoy))
OLS1 <- lm(cpi_geo_y ~ loc_geo_y)
par(mfrow=c(2,2))
plot(OLS1)

d + geom_smooth(method = lm, mapping = aes(x = gas_yoy, y = inflation_yoy)) + geom_point(mapping = aes(x = gas_yoy, y = inflation_yoy))
OLS2 <- lm(cpi_geo_y ~ gas_geo_y)
par(mfrow=c(2,2))
plot(OLS2)

d + geom_smooth(method = lm, mapping = aes(x = local_price_yoy, y = inflation_yoy)) + geom_point(mapping = aes(x = local_price_yoy, y = inflation_yoy))
OLS3 <- lm(cpi_geo_y ~ oil_geo_y)
par(mfrow=c(2,2))
plot(OLS3)

d + geom_smooth(method = lm, mapping = aes(x = gas_yoy, y = oil_yoy)) + geom_point(mapping = aes(x = gas_yoy, y = oil_yoy))
OLS4 <- lm(oil_geo_y ~ gas_geo_y)
par(mfrow=c(2,2))
plot(OLS4)
```



Individual Ordinary Least Squares

```{r pressure, echo=TRUE}


```
```{r}

```
```{r}

```
```{r}
OLS4 <- lm(oil_geo_y ~ gas_geo_y)
par(mfrow=c(2,2))
plot(OLS4)

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
